---
title: "Bio 304: Assignment XX -- Topic Name"
author: "Your Name and NetID here!"
output:
  html_document:
    highlight: default
    theme: readable
    fig_caption: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, prompt = TRUE, eval = TRUE, 
                      warning = TRUE, comment=NA, cache = FALSE,
                      fig.width = 6, fig.height = 4)
```


# Instructions

Create an R Markdown document and answer the questions below using code blocks that generate the correct outputs. Feel free to also use explanatory text in your markdown document, however **each of your solutions should show how to generate the answer using R code**.

Please name your file in the following manner:

* `Lastname_Firstinitial_Assignment0X.Rmd` -- For example: `Magwene_P_Assignment01.Rmd`

When completed, submit your assignment via the appropriate submission section on Sakai.

## Tips

* There should be **one code block per (sub)question**. A good rule of thumb is that each unit of code that produces a "stand alone" answer should be in its own code block.

* Format your code for readability.  This will help you debug your code and will help us grade your code, increasing the possibility you will get partial credit if you make a mistake somewhere along the line.

* Use headers to delineate different sections or questions.

* Libraries only need to be loaded once.

* Data structures and variables defined in one code block don't need to be repeated in subsequent code blocks (unless they've changed)

* **Knit your code** -- knitting your code creates a "clean" environment in which your code is evaluated and run, helping you to check that your notebook produces the answers you expect.


# Question 1

a. Question 1a: Repeat or summarize the question here.
 
```{r}
# your code here
```
  
Discussion of answer as needed.
 

b. Question 1b: Repeat or summarize the question here. 
 
```{r}
# your code here
```
  
Discussion of answer as needed.
  
# Question 2

a. Question 2a: Repeat or summarize the question here.
 
```{r}
# your code here
```
  
Discussion of answer as needed.
 

b. Question 2b: Repeat or summarize the question here. 
 
```{r}
# your code here
```
  
Discussion of answer as needed.

