---
title: "Bio 304: Assignment XX -- Topic Name"
author: "Your Name and NetID here!"
output:
  html_document:
    highlight: default
    theme: readable
    fig_caption: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, prompt = TRUE, eval = TRUE, 
                      warning = TRUE, comment=NA, cache = FALSE,
                      fig.width = 6, fig.height = 4)
```


# Instructions

Create an R Markdown document and answer the questions below using code blocks that generate the correct outputs. Feel free to also use explanatory text in your markdown document, however **each of your solutions should show how to generate the answer using R code**.

When completed, submit your assignment via the appropriate submission section on Sakai.


# Question 1

a. Question 1a
 
    ```{r}
    # your code here
    ```
  
    Discussion of answer as needed.
 

b. Question 1b 
 
    ```{r}
    # your code here
    ```
  
    Discussion of answer as needed.
  
# Question 2

a. Question 2a
 
    ```{r}
    # your code here
    ```
  
    Discussion of answer as needed.
 

b. Question 2b 
 
    ```{r}
    # your code here
    ```
  
    Discussion of answer as needed.

