---
title: "Confidence intervals"
author: "Paul M. Magwene"
output: beamer_presentation
fontsize: "10pt"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
options(width=50)
```



## Population of interest

```{r}
library(tidyverse)
set.seed(20171114)

height.data <- read_csv("https://github.com/bio304-class/bio304-course-notes/raw/master/datasets/simulated-US-male-height.csv")

true.values <- 
  height.data %>% 
  summarize(mean = mean(height), 
            sd = sd(height))


ggplot(height.data, aes(x = height)) +
  geom_histogram(aes(y = ..density..), bins=50, alpha=0.5) + 
  geom_density() +
  labs(x = "Height (cm)", title = "Distribution of Heights in the Population of Interest")
```

The population distribution of heights is $N(175.7, 15.19)$


## Point estimates of the mean for random samples of size 50

```{r, fig.height = 8, fig.width=3, fig.align = "center",  out.width="30%"}

rsample.stats <- function(x, n, var.name) {
  sample_x <- sample_n(x, n)
  data_frame(sample.size = n, 
             sample.mean = mean(sample_x[[var.name]]), 
             sample.sd = sd(sample_x[[var.name]]),
             sample.se = sample.sd/sqrt(n),
             ci.95.left = (sample.mean - 1.96 * sample.se),
             ci.95.right = (sample.mean + 1.96 * sample.se),
             ci.includes.mean = (ci.95.left <= true.values$mean) & (ci.95.right >= true.values$mean))
}

samples.of.50 <-
  rerun(1000,  rsample.stats(height.data, 50, "height")) %>%
  bind_rows()

samples <- samples.of.50[1:100,]

samples %>%
  ggplot(aes(x = sample.mean, y = seq(1,100))) + 
    geom_vline(xintercept = true.values$mean, color = "gray", alpha=0.75, linetype = "dashed", size = 1) +
    geom_point() +
    labs(x = "mean x", y = "sample", 
         title = "Point estimates of mean height\nfor 100 random samples of size 50")  
```



## 95% Confidence intervals for the mean for random samples of size 50

```{r, fig.height = 8, fig.width=4.5, fig.align = "center", out.width="40%"}

samples %>%
  ggplot(aes(x = sample.mean, y = seq(1,100), color = ci.includes.mean)) + 
    geom_vline(xintercept = true.values$mean, color = "gray", alpha=0.75, linetype = "dashed", size = 1) +
    geom_point() +
    geom_errorbarh(aes(xmin = ci.95.left, xmax = ci.95.right)) +
    scale_color_manual(values=c("red", "black")) +
    labs(x = "mean x and estimated CI", y = "sample", 
         title = "95% CI: mean ± 1.96×SE\nfor 100 samples of size 50")  
```


## Interpretting confidence intervals

From NIST page on confidence intervals:

> As a technical note, a 95 % confidence interval does not mean that there is a 95 % probability that the interval contains the true mean. The interval computed from a given sample either contains the true mean or it does not. Instead, **the level of confidence is associated with the method of calculating the interval** ... That is, for a 95% confidence interval, if many samples are collected and the confidence interval computed, in the long run about 95% of these intervals would contain the true mean.

## Confidence intervals: general formulation

We define the $(100\times\beta)$% confidence interval for the statistic $\phi$ as the interval:

\[
CI_\beta = \phi_{n} \pm (z \times {SE}_{\phi,n})
\]

Where:

* $\phi_{n}$ is the statistic of interest in a random sample of size $n$
* ${SE}_{\phi,n}$ is the standard error of the statistic $\phi$ (via simulation or analytical solution)

And the value of $z$ is chosen so that:

* across many different random samples of size $n$, the true value of the $\phi$ in the population of interest would fall within the interval approximately $(100\times\beta)$% of the time

