---
title: "Functions and Control Flow"
author: "Paul M. Magwene"
output: beamer_presentation
fontsize: "9pt"    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, comment = "", collapse=TRUE)
options(width=60)
```



## Functions

* Functions are a programming tool for organizing and abstract related computations.  

* They help to logically structure analyses and programs.

\bigskip

To write your own function in R, define a function like this:

\medskip

```{r, echo=TRUE}
func.name <- function(arg) {
  # one or more statements 
  # return final result (if any) as last line
}
```

## Example functions


```{r}
add.two <- function(x) {
  return(x + 2)
}
```

The function `add.two`:

  - takes a single argument, labelled `x`
  - returns the value `x + 2` 
  

## `return()` is optional

```{r}
add.two <- function(x) {
  x + 2
}

rev.seq <- function(x){
  rev(seq(x))
}

```

- The value of the last expression is automatically returned so simple functions don't need a return statement



## Arguments are placeholders for input


```{r}
add.two <- function(x) {
  x + 2
}

x <- 5
add.two(3)  # returns 6 not 10
```

## Functions can take multiple arguments

```{r}
squared.difference <- function(x, y) {
  (x - y)**2
}

squared.difference(10, 5)
```

## Arguments can have default values

```{r}
squared.difference <- function(x, y=0) {
  (x - y)**2
}


squared.difference(10)      # y defaults to zero
squared.difference(10, 5)   # y is 5, inferred by position
squared.difference(10, y=5) # y is explcitily set as 5
```
